---
title: "Coursera Practical Machine Learning Project"
author: "Ezeonyebuchi E. C."
date: "27/04/2019"
output:
  html_document:
    keep_md: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Introduction:
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount 
of data about personal activity relatively inexpensively. These type of devices are part of the quantified 
self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health,
to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

In this project, I will use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participant 
who were asked to perform dumbbell lifts correctly and incorrectly in 5 different ways.
The data consists of a Training data and a Test data (to be used to validate the selected model).
The goal of your project is to predict the manner in which they did the exercise. 
This is the “classe” variable in the training set. I will useany of the other variables to predict with.

Load needed libraries
```{r}
library(knitr)
library(rattle)
library(randomForest)
library(rpart.plot)
library(caret)
library(rpart)
library(gbm)
library(RColorBrewer)
library(corrplot)
```
Downloading the websites
1).
For the traing data
```{r}
TrainURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
##download.file(TrainURL, destfile = "./pml-training", method="auto")
```

2).For test data
```{r}
TestUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
##download.file(TestUrl, destfile = "./pml-testing.csv", method = "auto")
```
Reading, Cleaning and Exploring the data
```{r}
trainData <- read.csv("./pml-training", header = TRUE)
#head(trainData)
dim(trainData)
#str(trainData)
#summary(trainData)

testData <- read.csv("./pml-testing.csv", header = TRUE)
#head(testData)
dim(testData)
#str(testData)
#summary(testData)
```
Partioning the training set into two
Partioning Training data set into two data sets.I am using 70% for Training and 30% for Testing:

```{r}
inTrain <- createDataPartition(y = trainData$classe, p=0.7, list=FALSE)
Trainset <- trainData[inTrain, ]
Testset <- trainData[-inTrain, ]
dim(Trainset)
dim(Testset)
```
Removing variables with Nearly Zero Variance
```{r}
NZV <- nearZeroVar(Trainset)
Trainset <- Trainset[, -NZV]
Testset  <- Testset[, -NZV]
dim(Trainset)
dim(Testset)
```
Remove variables that are mostly NA (mNA)
```{r}
mNA <- sapply(Trainset, function(x) mean(is.na(x))) > 0.95
Trainset <- Trainset[, mNA==FALSE]
Testset  <- Testset[, mNA==FALSE]
dim(Trainset)
dim(Testset)
```
Remove identification only variables (columns 1 to 5)
```{r}
Trainset <- Trainset[, -(1:5)]
Testset  <- Testset[, -(1:5)]
dim(Trainset)
dim(Testset)
```
The number of variables for analysis has been reduced to 54 with the above cleaning exercise 
Correlation Analysis
```{r}
corMatrix <- cor(Trainset[, -54])
corrplot(corMatrix, order = "FPC", method = "color", type = "lower", 
         tl.cex = 0.8, tl.col = rgb(0, 0, 0))
```
Prediction Models Building
Three methods will be applied to the regression models. 
A Confusion Matrix is plotted at the end of each for better visualization of the  the accuracy of the models.

(a)Random Forest
model fit
```{r}
controlRF <- trainControl(method="cv", number=3, verbose=FALSE)
modFitRandForest <- train(classe ~ ., data=Trainset, method="rf",
                          trControl=controlRF)
modFitRandForest$finalModel
```
Prediction on Test Dataset
```{r}
predictRandForest <- predict(modFitRandForest, newdata=Testset)
confMatRandForest <- confusionMatrix(predictRandForest, Testset$classe)
confMatRandForest

plot(confMatRandForest$table, col = confMatRandForest$byClass, 
     main = paste("Random Forest - Accuracy =",
                  round(confMatRandForest$overall['Accuracy'], 4)))
```
(b) Method: Decision Trees
model fit
```{r}
set.seed(4000)
modFitDT <- rpart(classe ~ ., data=Trainset, method="class")
fancyRpartPlot(modFitDT)
```
Prediction on Test dataset
```{r}
predictDT <- predict(modFitDT, newdata=Testset, type="class")
confMatDT <- confusionMatrix(predictDT, Testset$classe)
confMatDT

plot(confMatDT$table, col = confMatDT$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(confMatDT$overall['Accuracy'], 4)))
```

(c)Generalized Boosted Model
model fit
```{r}
set.seed(4000)
control_GBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFit_GBM  <- train(classe ~ ., data=Trainset, method = "gbm",
                     trControl = control_GBM, verbose = FALSE)
modFit_GBM$finalModel
```
Prediction on Test dataset
```{r}
predict_GBM <- predict(modFit_GBM, newdata=Testset)
confMat_GBM <- confusionMatrix(predict_GBM, Testset$classe)
confMat_GBM


plot(confMat_GBM$table, col = confMat_GBM$byClass, 
     main = paste("GBM - Accuracy =", round(confMat_GBM$overall['Accuracy'], 4)))
```
The accuracy of the 3 regression modeling methods above are:
```{r}
RandomForest <- print(paste("Random Forest - Accuracy =",
                            round(confMatRandForest$overall['Accuracy'], 4)))

DecisionTree <- print(paste("Decision Tree - Accuracy =",
                            round(confMatDT$overall['Accuracy'], 4)))

GBM <- print(paste("GBM - Accuracy =", round(confMat_GBM$overall['Accuracy'], 4)))
```
From the above observation, RandomForest model gives a better result and will be applied to predict the 20 quiz results 
(testing (testData) dataset) as shown below.
```{r}
predictTEST <- predict(modFitRandForest, newdata=testData)
predictTEST
```